---
title: "DATA606_Project_Proposal_RJM"
author: "RJM"
date: '2020-02-13'
always_allow_html: true
output:
  
  pdf_document: default
  word_document: default
  html_document: default
---


### Data Preparation

The following is the code for the data preparation:

```{r global}
library(flexdashboard)
library(DT)
library(ggplot2)
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
```


```{r DatasetImport}

rm(list = ls())

#disable scientific notation, so that actual decimal values are imported instead of exponential factors
options(scipen = 999)


# Importing country Metadta dataset into R
download.file("https://github.com/rjmirza/DATA-606/raw/master/final_project/datasets/GDP.xls", "GDP.xls")
country_metadata_dataset <- read_excel("GDP.xls", col_names = TRUE, sheet = "Metadata - Countries")

# Importing GDP (1995-2018) by country dataset into R
gdp_dataset <- read_excel("GDP.xls", col_names = TRUE, sheet = "Data", skip = 3) %>%
  data.frame(., stringsAsFactors = F) %>%
  select(., 1,2,3,40:63)

# Importing GDP percapita (1995-2018) by country dataset into R
download.file("https://github.com/rjmirza/DATA-606/raw/master/final_project/datasets/GDP%20per%20Capita.xls", "GDP_per_Capita.xls")
gdp_percapita_dataset <- read_excel("GDP_per_Capita.xls", col_names = TRUE, sheet = "Data", skip = 3) %>%
  data.frame(., stringsAsFactors = F) %>%
  select(., 1,2,3,40:63)

# Importing Manufacturing GDP (1995-2018) percentage by country dataset into R
download.file("https://github.com/rjmirza/DATA-606/raw/master/final_project/datasets/Manufacturing.xls", "Manufacturing.xls")
gdp_manufacturing_dataset <- read_excel("Manufacturing.xls", col_names = TRUE, sheet = "Data", skip = 3) %>%
  data.frame(., stringsAsFactors = F) %>%
  select(., 1,2,3,40:63)

# Importing Agriculture GDP (1995-2018) percentage by country dataset into R
download.file("https://github.com/rjmirza/DATA-606/raw/master/final_project/datasets/Agriculture.xls", "Agriculture.xls")
gdp_agriculture_dataset <- read_excel("Agriculture.xls", col_names = TRUE, sheet = "Data", skip = 3) %>%
  data.frame(., stringsAsFactors = F) %>%
  select(., 1,2,3,40:63)

# Importing Service GDP (1995-2018) percentage by country dataset into R
download.file("https://github.com/rjmirza/DATA-606/raw/master/final_project/datasets/Service.xls", "Service.xls")
gdp_service_dataset <- read_excel("Service.xls", col_names = TRUE, sheet = "Data", skip = 3) %>%
  data.frame(., stringsAsFactors = F) %>%
  select(., 1,2,3,40:63)

# Importing Industries GDP (1995-2018) percentage by country dataset into R
download.file("https://github.com/rjmirza/DATA-606/raw/master/final_project/datasets/Industries.xls", "Industries.xls")
gdp_industries_dataset <- read_excel("Industries.xls", col_names = TRUE, sheet = "Data", skip = 3) %>%
  data.frame(., stringsAsFactors = F) %>%
  select(., 1,2,3,40:63)

# Importing Ores_Metals_Minerals GDP (1995-2018) percentage by country dataset into R
download.file("https://github.com/rjmirza/DATA-606/raw/master/final_project/datasets/Ores_Metals_Minerals.xls", "Ores_Metals_Minerals.xls")
gdp_ores_metals_minerals_dataset <- read_excel("Ores_Metals_Minerals.xls", col_names = TRUE, sheet = "Data", skip = 3) %>%
  data.frame(., stringsAsFactors = F) %>%
  select(., 1,2,3,40:63)

```

```{r MergeDF}

df1 <- gather(gdp_dataset, "year", "GDP", 4:27) %>% select(1, 4, 5)
df1$GDP <- df1$GDP/1000000
df2 <- gather(gdp_percapita_dataset, "year", "GDP Percapita", 4:27) %>% select(1, 4, 5)
df3 <- gather(gdp_industries_dataset, "year", "Industry Percent of GDP", 4:27) %>% select(1, 4, 5)
df4 <- gather(gdp_service_dataset, "year", "Services Percent of GDP", 4:27) %>% select(1, 4, 5)
df5 <- gather(gdp_agriculture_dataset, "year", "Agriculture Percent of GDP", 4:27) %>% select(1, 4, 5)
df6 <- gather(gdp_manufacturing_dataset, "year", "Manufacturing Percent of GDP", 4:27) %>% select(1, 4, 5)
df7 <- gather(gdp_ores_metals_minerals_dataset, "year", "Ores_Metals_Minerals Percent of GDP", 4:27) %>% select(1, 4, 5)
df <- merge(df1, df2, all.y = T)
df <- merge(df, df3, all.y = T)
df <- merge(df, df4, all.y = T)
df <- merge(df, df5, all.y = T)
df <- merge(df, df6, all.y = T)
df <- merge(df, df7, all.y = T)
df <- merge(country_metadata_dataset, df, by.x = "TableName", by.y = "Country.Name", all.y = T)
summary(df)
# removing characters from the year and converting the type to numeric
df$year <- str_extract(df$year, "[:digit:]+") %>%
  as.numeric(df$year)
```

```{r IncomeGroupsDF}
incomegroup_df <- df %>%
  filter(., is.na(IncomeGroup)) %>%
  filter(., `Country Code` %in% c("EAR","FCS","HIC","HPC","LDC","LIC","LMC","LMY","LTE","MIC","PRE","PST","UMC")) %>%
  arrange(TableName, year)
```

```{r RegionsDF}
economy_by_region_df <- df %>%
  filter(., is.na(IncomeGroup)) %>%
  filter(., TableName %in% c("East Asia & Pacific","Europe & Central Asia","Latin America & Caribbean","Middle East & North Africa","North America","South Asia","Sub-Saharan Africa")) %>%
  select(1,3,4,6:8,10:13) %>%
  arrange(TableName, year)

```
Input Datasets and import
=======================================================================

 - GDP(GDP.xls)                         : GDP by country
 - GDP Percapita(GDP_per_Capita.xls)    : GDP Percapita by country
 - Manufacturing GDP(Manufacturing.xls) : Manufacturing GDP percentage by country
 - Agriculture GDP(Agriculture.xls)     : Agriculture GDP percentage by country
 - Service GDP(Service.xls)             : Service GDP percentage by country
 - Industries GDP(Industries.xls)       : Industries GDP percentage by country
 - Ores_Metals_Minerals GDP(Ores_Metals_Minerals.xls)       : Ores_Metals_Minerals GDP percentage by country


### Research question 

The question chosen to answer was to figure out whether there is a relationship between a country's wealth and the primary type of industry that its economy is mainly engaged in. In other words, the aim was to find the impact of main industries in particular countries or regions on the economy and general well-being of its inhabitants. To approach it from a research angle, we used the Worldbank's datasets to find the main industries in the countries or regions. The two main industry categories we used were knowledge-based and traditional. The knowledge-based industries was further segregated into manufacturing and services while agriculture and minerals were condidered to be sub-categories for traditional industries.

```{r KnowledgeTraditinoalDF}
knowledge_traditinoal_dF <- df %>%
  filter(., !is.na(IncomeGroup)) %>%
  select(1,3,4,6,7,10:13) %>%
  mutate("Knowledge based Percent of GDP" = ifelse(is.na(`Services Percent of GDP`), 0, `Services Percent of GDP`)+
           ifelse(is.na(`Manufacturing Percent of GDP`),0,`Manufacturing Percent of GDP`),
         "Traditinoal based Percent of GDP" = ifelse(is.na(`Agriculture Percent of GDP`),0,`Agriculture Percent of GDP`)+
           ifelse(is.na(`Ores_Metals_Minerals Percent of GDP`),0,`Ores_Metals_Minerals Percent of GDP`)) %>%
  arrange(TableName, year)

country_gdp_mean_sd_dF <- knowledge_traditinoal_dF %>%
  group_by(TableName) %>%
  summarise("Country Mean GDP" = mean(GDP, na.rm=TRUE),
            "Country SD GDP" = sd(GDP, na.rm=TRUE)
            )

world_knowledge_gdp_percent_mean_dF <- knowledge_traditinoal_dF %>%
  group_by(year) %>%
  summarise("World Mean Knowledge GDP percent" = mean(`Knowledge based Percent of GDP`, na.rm=TRUE))

world_traditional_gdp_percent_mean_dF <- knowledge_traditinoal_dF %>%
  group_by(year) %>%
  summarise("World Mean Traditinoal GDP percent" = mean(`Traditinoal based Percent of GDP`, na.rm=TRUE))

knowledge_traditinoal_dF <- knowledge_traditinoal_dF %>%
  merge(., country_gdp_mean_sd_dF, by.x = "TableName", by.y = "TableName", all.y = T) %>%
  mutate("Country SD GDP in percent" = `Country SD GDP`/`Country Mean GDP`*100) %>%
  merge(., world_knowledge_gdp_percent_mean_dF, by.x = "year", by.y = "year", all.y = T) %>%
  merge(., world_traditional_gdp_percent_mean_dF, by.x = "year", by.y = "year", all.y = T) %>%
  select(1:5,10:16) %>%
  na_if(., 0) %>%
  arrange(TableName, year)

```

### Cases 

There were 263 cases representing each country and independent territory. The details on these cases form the bulk of data.


### Data collection 

We thought that the industry composition of GDP is an indicator that will tell us about the main industries operating in a certain country. We used Worldbank's website to collect data on the four catrgories namely agriculture, minerals, services, and manufacturing. The data had to be curated to fit it between the years of 1995-2018 as most of the countries had missing information for years before 1995. We had to nuance the data analysis by introducing factors like recessions, per capita income, and regions.

```{r}
### initial dataframe
DT::datatable(df, options = list(pageLength = 5))

### incomegroup dataframe
DT::datatable(incomegroup_df, options = list(pageLength = 5))

### economy by region dataframe
DT::datatable(economy_by_region_df, options = list(pageLength = 5))

### knowledge and traditinoal GDP's dataframe
DT::datatable(knowledge_traditinoal_dF, options = list(pageLength = 5))
```


### Type of study 

This is an observational study where we took existing data to support our hypothesis.


### Data Source 

The sources are as follows:

1. GDP

https://data.worldbank.org/indicator/NY.GDP.MKTP.CD

2. GDP per Capita

https://data.worldbank.org/indicator/NY.GDP.PCAP.CD

3. Agricultrure, forestry, and fishing as % of GDP

https://data.worldbank.org/indicator/NV.AGR.TOTL.ZS

4. Ores and Metals exports (% of merchandise exports) taken as a proxy to mineral production and exports

https://data.worldbank.org/indicator/TX.VAL.MMTL.ZS.UN

5. Service, value added (% of GDP)

https://data.worldbank.org/indicator/NV.SRV.TOTL.ZS

6. Manufacturing, value added (% of GDP)

https://data.worldbank.org/indicator/NV.IND.MANF.ZS



### Dependent Variable


The response variable is the GDP output which is quantitative measured in either percentage points or dollar amounts.


### Independent Variable


The qualitative independent variables are the countries, regions, and types of main industries these places are engaged in. The quantitative independent variable is the years to measure the changes in the GDP output.

### Relevant summary statistics 

```{r}
NIG <- length(unique(incomegroup_df[["TableName"]]))

valueBox(NIG, color = "primary")
```

### GDP in millions by income groups

There were some challenges with imputing the missing data. We figured that leaving them blank is the best to get the continuous lines on graphs.
```{r fig.width=10}

ggplot(incomegroup_df, aes(x=factor(year), colour=TableName, group = TableName)) +
  geom_point(aes(y = `GDP`)) + 
  geom_line(aes(y = `GDP`)) +
  theme(axis.text.x = element_text(size=10, angle=90)) +
  theme(axis.text.y = element_text(size=10, angle=90)) +
  labs(title = "GDP by income groups in millions")
```

For more information on demographic dividend, please refer to the following link:

https://www.imf.org/external/pubs/ft/fandd/2006/09/basics.htm


### GDP in million based on specific geological regions

The regions helped us to gain insights through pre-existing knowledge regarding the wealth and the main industries they indulge into. For example, we know that Europe is a Service and Manufacturing intensive economy and is lumped with Central Asia which is a Minerals dependent industry. Data could be refined further to separate these regions and find out how much of the GDP is dependent on the specific industries!
```{r}

library(ggplot2)
ggplot(economy_by_region_df, aes(x=factor(year), colour=TableName, group = TableName)) +
  geom_point(aes(y = `GDP`)) + 
  geom_line(aes(y = `GDP`)) +
  theme(axis.text.x = element_text(size=10, angle=90)) +
  theme(axis.text.y = element_text(size=10, angle=90)) +
  labs(title = "GDP by Region in millions")

```

### Services Percent of GDP

Services make a high portion of GDP in High income countries.

```{r fig.width=10}
ggplot(incomegroup_df, aes(x=factor(year), colour=TableName, group = TableName)) +
  geom_point(aes(y = `Services Percent of GDP`)) + 
  geom_line(aes(y = `Services Percent of GDP`)) +
  theme(axis.text.x = element_text(size=10, angle=90)) +
  theme(axis.text.y = element_text(size=10, angle=90)) +
  labs(title = "Services Percent of GDP by income groups")
```


### Agriculture Percent of GDP

The Agriculture seems to be trending downward in every type of economy. An interesting discovery would be to see if it is losing share to other industries or it is declining as an industry!

```{r fig.width=10}
ggplot(incomegroup_df, aes(x=year, colour=TableName, group = TableName)) +
  geom_point(aes(y = `Agriculture Percent of GDP`)) + 
  geom_line(aes(y = `Agriculture Percent of GDP`)) +
  theme(axis.text.x = element_text(size=10, angle=90)) +
  theme(axis.text.y = element_text(size=10, angle=90)) +
  labs(title = "Agriculture Percent of GDP by income groups")
```